The Leaky Semicolon

### 2021-10-27 paper/related.tex
removing todos
updated related wokr

There are few prior studies of relaxed memory that include sequential
composition and/or precise calculation of semantic dependencies.
\citet{DBLP:journals/pacmpl/JagadeesanJR20} give a denotational semantics,
using prefixing rather than sequential compositions.
\citet{DBLP:conf/esop/PaviottiCPWOB20} give a denotational semantics,
calculating dependencies using event structures rather than logic.  They give
the semantics of sequential composition in continuation passing style,
whereas we give it in direct style.  This paper provides a general technique
for computing sequential dependencies and applies it to these two approaches.

There are many differences between \PwP{}
\cite{DBLP:journals/pacmpl/JagadeesanJR20} and \PwT{}.  Two of these are
motivated by requirements unique to \PwT{}: First, \PwP{} uses substitution
to define read actions; \PwT{} Skolemizes in order to satisfy disjunction
closure for predicate transformers (\textsection\ref{sec:ex:assoc}).  Second,
\PwP{} requires \emph{consistency} and \emph{causal strengthening}; \PwT{}
drops these in order to satisfy associativity (\reflem{lem:monoid}).  Other
differences are stylistic: For example, we replaced the termination
\emph{actions} of \PwP{} by termination \emph{conditions}---our formulation
fixes an error in \citeauthor{DBLP:journals/pacmpl/JagadeesanJR20}'s
definition of parallel composition.  We also fix an error in their treatment
of redundant read elimination.

@@

Other studies of relaxed memory can be categorized by their approach to
dependency calculation.  Hardware models use syntactic dependencies
\cite{alglave}.  Many software models do not bother with dependencies at all
\cite{Batty:2011:MCC:1926385.1926394, DBLP:journals/pacmpl/WattRP19,
  DBLP:conf/pldi/WattPPBDFPG20, goMM}.  Others have strong dependencies that
disallow compiler optimizations and efficient implementation, typically
requiring fences for every relaxed read on Arm
\cite{Lamport:1979:MMC:1311099.1311750, DBLP:conf/pldi/LahavVKHD17,
  Dolan:2018:BDR:3192366.3192421, DBLP:conf/lics/JeffreyR16,
  Boehm:2014:OGA:2618128.2618134}. %, DBLP:journals/corr/abs-1804-04214}.
%
Many of the most prominent models are operational models based
on speculative execution \cite{Manson:2005:JMM:1047659.1040336,
  DBLP:conf/esop/JagadeesanPR10,
  DBLP:conf/popl/KangHLVD17,DBLP:journals/pacmpl/ChakrabortyV19,DBLP:conf/pldi/LeeCPCHLV20,promising-ldrf}.
Morally, \PwT{} fits between the strong models and the speculative ones.
% We provide a detailed comparison with these approaches in \textsection\ref{sec:promising}.

>>>>

There are few prior studies of relaxed memory that include sequential
composition and/or precise calculation of semantic dependencies.  No prior
work gives the semantics of sequential composition in direct style.
\citet{DBLP:conf/esop/PaviottiCPWOB20} defined \MRD{}, which calculates
dependencies using event structures rather than logic.  This strategy is
brittler than ours, leading to false positives (\textsection\ref{sec:lir}).
\citet{DBLP:journals/pacmpl/JagadeesanJR20} defined \PwP{}, using logical
entailment to define dependency.  Although \PwT{} is based on \PwP{}, there
are many differences.  Two of these are motivated by requirements unique to
\PwT{}: First, \PwP{} uses substitution to define read actions; \PwT{}
Skolemizes in order to satisfy disjunction closure for predicate transformers
(\textsection\ref{sec:ex:assoc}).  Second, \PwP{} requires \emph{consistency}
and \emph{causal strengthening}; \PwT{} drops these in order to satisfy
associativity (\reflem{lem:monoid}).  Other differences are stylistic: For
example, we use termination \emph{conditions} rather than termination
\emph{actions}---our formulation fixes an error in
\citeauthor{DBLP:journals/pacmpl/JagadeesanJR20}'s definition of parallel
composition.  We also fix an error in their treatment of redundant read
elimination (\textsection\ref{sec:semreg}).

@@

Other studies of relaxed memory can be categorized by their approach to
dependency calculation.  Hardware models use syntactic dependencies
\cite{alglave}.  Many software models do not bother with dependencies at all
\cite{Batty:2011:MCC:1926385.1926394, DBLP:journals/pacmpl/WattRP19,
  DBLP:conf/pldi/WattPPBDFPG20, goMM}.  Others have strong dependencies that
disallow compiler optimizations and efficient implementation, typically
requiring fences for every relaxed read on Arm
\cite{Lamport:1979:MMC:1311099.1311750, DBLP:conf/pldi/LahavVKHD17,
  Dolan:2018:BDR:3192366.3192421, DBLP:conf/lics/JeffreyR16,
  Boehm:2014:OGA:2618128.2618134}. %, DBLP:journals/corr/abs-1804-04214}.
%
Many of the most prominent models are operational models based on speculative
execution \cite{Manson:2005:JMM:1047659.1040336,
  DBLP:conf/esop/JagadeesanPR10,
  DBLP:conf/popl/KangHLVD17,DBLP:journals/pacmpl/ChakrabortyV19,DBLP:conf/pldi/LeeCPCHLV20,promising-ldrf}.

Morally, \PwT{} fits between the strong models and the speculative ones.
Looking at the details, however, \PwTmca{} is incomparable to both \rcXI{}
\cite{DBLP:conf/pldi/LahavVKHD17} and the promising semantics
\cite{DBLP:conf/popl/KangHLVD17}, to take two examples.  \rcXI{} allows
non-\mca{} behaviors that \PwTmca{} disallows.  \PwTmca{} has a weaker notion
of coherence than the promising semantics.

% We provide a detailed comparison with these approaches in \textsection\ref{sec:promising}.
